---
title: "Homework 2"
author: Yali Zhai
output: github_document
---


```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data_hw2/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
 
 
```

read precipitation for 2018 and 2017.

```{r}
precip_2018 = 
  read_excel(
    "./data_hw2/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data_hw2/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annul precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

The dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kind of trash. There are a total od `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month percipitation data. The total precipitation in 2018 is `r sum(pull(precip_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(trashwheel_df, sports_balls))`.

## Problem 2

Read the NYC Transit data

```{r}
transit_data = 
  read_csv(
    "./data_hw2/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") 

transit_df =  
  janitor::clean_names(transit_data) %>%
  select("line", "station_name", "station_latitude", "station_longitude", "route1", "route2", "route3", "route4", "route4", "route5", "route6", "route7", "route8", "route9", "route10", "route11", "entrance_type", "entry", "vending", "ada") %>% 
  mutate(entry = recode(entry,"YES" = TRUE, "NO" = FALSE ))

```

The dataset contains `r nrow(transit_data)` observations and `r ncol(transit_data)` variables. The variables are `r names(transit_data)`.
So far, I have done some clean work. First, I change the variable names to lowercase. Then I choose the columns that we need. Lastly, I convert the entry variable from character to a logical variable. The dimension of the resulting dataset is `r nrow(transit_df)` * `r ncol(transit_df)`. These data are not tidy enough.

```{r}
distinct_stat = distinct(transit_df, line, station_name)

length(which(pull(transit_df, ada) == TRUE))

no_vending_df = filter(transit_df, vending == "NO")
nrow(no_vending_df)
length(which(pull(no_vending_df,entry) == TRUE))
prop = length(which(pull(no_vending_df,entry) == TRUE)) / nrow(no_vending_df)


```

There are `r nrow(distinct_stat)` distinct stations. `r length(which(pull(transit_df, ada) == TRUE))` stations are ADA compliant.
The proportion of station entrances / exits without vending allow entrance is `r prop`.


Reformat data

```{r}
transit_df = 
  transit_df %>% 
  mutate_at(vars(route8:route11), as.character)

transit_tidy_data = 
  pivot_longer(
    transit_df,
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )

distinc_tidy_df = distinct(transit_tidy_data, line, station_name, .keep_all = TRUE)

route_serve_A = filter(distinc_tidy_df, route_name == "A")
nrow(route_serve_A)
length(which(pull(route_serve_A, ada) == TRUE))
```

There are `r nrow(route_serve_A)` distinct stations serve the A train. Among these stations, there are `r length(which(pull(route_serve_A, ada) == TRUE))` are ADA compliant.


## Problem 3

Read pols-month dataset.

```{r}
pols_month_df = 
  read_csv("./data_hw2/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = recode(month, `01` = "January", `02` = "Feburary", `03` = "March", `04` = "April", `05` = "May", `06` = "June", `07` = "July", `08` = "August", `09` = "September", `10` = "October", `11` = "November", `12` = "December"),
    president = ifelse(prez_gop == 1, "gop", "dem")
  ) %>% 
  select(-day, -prez_gop, -prez_dem)
```


Read the snp data.

```{r}
snp_df = 
  read_csv("./data_hw2/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(
    month = recode(month, `1` = "January", `2` = "Feburary", `3` = "March", `4` = "April", `5` = "May", `6` = "June", `7` = "July", `8` = "August", `9` = "September", `10` = "October", `11` = "November", `12` = "December")) %>% 
  select(-day) %>% 
  relocate(year, month)

```

Read the unemployment data.

```{r}
unemploy_df = 
  read.csv("./data_hw2/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemplyment"
  ) %>% 
  mutate(
    month = month.name[match(month, str_to_lower(month.abb))],
    year = as.character(year)
    )

```

